---
title: "RSI,GARD Cohort-Analysis"
output:
  html_document:
    df_print: paged
  rmdformats: html_clean
editor_options:
  chunk_output_type: inline
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo=TRUE, cache=FALSE, prompt=FALSE,
                      comment=NA,message=FALSE,warning=FALSE)



```


```{r load_packages}

library(gridExtra)
library(grid)
library(ggstance)
library(colorspace)
library(tidyverse)
library(dplyr)

library(readxl)
library(survival)
library(survminer)

library(rms)
library(knitr)
library(rmdformats)
library(xtable)

library(here)


options(xtable.floating = FALSE, xtable.timestamp = "", xtable.comment = FALSE)
options(max.print="75")
opts_knit$set(width=75)


```


# Load Data

```{r load_data}
rsi.all<-read.csv(file = "Data/data.csv")
rsi.all$Site<-as.character(rsi.all$Site)

```


```{r, include = FALSE}

# load color meta

q <- c("#2980b9","#10ac84","#8e44ad","#7f8c8d","#34495e","#b71c1c",rgb(.9,.55,.2),"#fb8c00","#27ae60","#4f4f4f","#2b2b2b")
q2 <-        c("#ff5252","#34ace0",   "#33d9b2","#706fd3","#ff793f","#3867d6",  "#808e9b","#f7b731","#2b2b2b")
names(q2) <- c("Breast", "Endometrium","TNBC", "Pancreas", "Head & Neck","NSCLC", "Glioma","Melanoma","Pooled")
q3 <- c(
        # "#ff0011",
        "#ff5252",
        "#f071ac", # "#ff526f", #"#ff1414",
        "#c92c2c",
        "#34ace0", 
        "#33d9b2", "#077d61",
        "#706fd3",
        "#ff793f",
        "#3867d6", 
        "#808e9b",
        "#f7b731",
        "#2b2b2b")

names(q3) <- c("Breast (Erasmus)" ,   "Breast (Karolinksa)" , "Breast (NKI)"  ,
   "Endometrium (TCC)" ,
   "TNBC (NKI)" ,  "TNBC (MCC)"  ,
   "Pancreas (TCC)"  ,
   "Head & Neck (NKI)" ,
   "NSCLC (MCC)" ,
   "Glioma (TCGA)" , 
  "Melanoma (TCC)" ,
  "Pooled")

```

# Clean data

## Calculate GARD 

```{r calc_GARD}
rsi.all<-add_column(rsi.all, 'alpha' = NA)
rsi.all<-add_column(rsi.all, 'GARD'  = NA)
rsi.all<-rsi.all %>% unite('Source_Site',c(Source,Site), sep="_", remove = F)


# sham doses
rsi.all<-rsi.all %>% 
  mutate(d = if_else(Received_RT==0, 2, d)) %>%
  mutate(TD = case_when(
  Received_RT==0 & Site == 'breast' ~ 50,
  Received_RT==0 & Site == 'endometrial' ~ 54,
  Received_RT==0 & Site == 'glioma' ~ 60,
  Received_RT==0 & Site == 'pancreas' ~ 50,
  Received_RT==0 & Site == 'melanoma' ~ 60,
  Received_RT==1 ~ TD
  )) %>%
  mutate(n = if_else(Received_RT==0,TD/2,n))

rsi.all <- rsi.all %>%
  mutate(alpha=log(RSI)/(-2) - 0.05*2) %>%
  mutate(GARD=n*d*(alpha + 0.05*d))

rsi.all <- rsi.all %>%
  mutate(alpha.beta = case_when(
    Site == 'breast' ~ 10,
    Site == 'endometrial' ~ 10,
    Site == 'lung' ~ 10,
    Site == 'glioma' ~ 10,
    Site == 'pancreas' ~ 10,
    Site == 'breast_TN' ~ 10,
    TRUE ~ 10
  )) %>%
  mutate(
    EQD2 = TD*( (d + alpha.beta)/(2 + alpha.beta) )
  )
```


## Adjust citations, cite, and cohort labels
```{r add_sources}
rsi.all<-rsi.all %>% mutate(Source = recode(Source, NKI_wboost='NKI',MCC_wboost='MCC', Pranama="NKI"))

rsi.all<-rsi.all %>% 
  mutate(Citation = case_when(
    Site=='endometrial' ~ "Fenstermacher2011",
    Source=="Karolinksa" ~ "Pawitan2005",
    Source=='Erasmus' ~ "Wang2005,Yu2007",
    Site=="pancreas" ~ "Fenstermacher2011",
    Site=='glioma' ~ "TCGA",
    Source=='NKI' & (Site=='breast_TN' | Site=='breast') ~ "Servant2012",
    Source=='MCC' & Site =='breast_TN' ~ "Fenstermacher2011",
    Site=='melanoma' ~ "Strom2017",
    Site=="HN" ~ "Pramana2007",
    Site=="lung" ~ "Fenstermacher2011"
  ))%>%
  mutate(Site_corrected = recode(Site, breast='Breast',  breast_TN='TNBC', melanoma='Melanoma',
                         endometrial='Endometrium', glioma = 'Glioma',
                         HN='Head & Neck', lung='NSCLC',pancreas='Pancreas')) %>%
  mutate(Cohort = sprintf("%s (%s)",Site_corrected,Source))

```


## Separate groups of interest

```{r sep_groups}
rt.all <- filter(rsi.all, Received_RT==1)
  
rt.rec <- rt.all %>%
  filter(!is.na(Event)) %>%
  select(Source_Site, Source, Site, Cohort, Citation,  RSI, GARD, Time, Event, TD, EQD2)

rt.os <- rt.all %>%
  filter(!is.na(Event_OS)) %>%
  select(Source_Site, Source, Site, Cohort, Citation,  RSI, GARD, Time_OS, Event_OS, TD, EQD2, MGMT_Expression)

noRT.all <- rsi.all %>% 
  filter(Received_RT==0)
  
noRT.rec <- noRT.all %>%
  filter(!is.na(Event)) %>%
  select(Source_Site, Source, Site, Cohort, Citation,  RSI, GARD, Time, Event, TD, EQD2)

noRT.os <- noRT.all %>%
  filter(!is.na(Event_OS)) %>%
  select(Source_Site, Source, Site, Cohort, Citation,  RSI, GARD, Time_OS, Event_OS, TD, EQD2, MGMT_Expression)

all.rec <- rsi.all %>%
  filter(!is.na(Event)) %>%
  select(Source_Site, Source, Site, Cohort, Citation,  Received_RT, RSI, GARD, Time, Event, TD, EQD2)

all.os <- rsi.all %>% 
  filter(!is.na(Event_OS)) %>%
  select(Source_Site, Source, Cohort, Citation,  Site, Received_RT, RSI, GARD, Time_OS, Event_OS, TD, EQD2, MGMT_Expression)
```


# Data Visualization

## GARD distribution between cohorts

### Recurrence Cohorts

```{r violin_GARD_rec}

ggplot(rt.rec, aes(x = GARD, y = Cohort,fill = Cohort)) + 
  geom_violin(color = "black") + 
  geom_boxplot(fill = "white", width = 0.15) +
  ggtitle("Distribution of GARD by Recurrence Cohorts")
  

```

### Survival Cohorts


```{r violin_GARD_OS}

ggplot(rt.os, aes(x = GARD, y = Cohort,fill = Cohort)) + 
  geom_violin(color = "black") + 
  geom_boxplot(fill = "white", width = 0.15) +
  ggtitle("Distribution of GARD by OS Cohorts")

```

## KM with all cohorts

We'll start with previous code written by Geoffrey Sedor. 

### Recurrence

```{r, fig.align="center"}
km.df <- all.rec 
km.rec <- survfit(Surv(Time,Event)~ Cohort + Received_RT, data = km.df)

for (i in 1:length(km.rec$strata)) {
  cur_strata = names(km.rec$strata[i])
  count = unname(km.rec$strata[i])
  c <- c(rep(cur_strata,count))
  if (i==1){
    c_all.rec <- c
  } else {
    c_all.rec <- c(c_all.rec,c)
  }
}

km.rec_tbl <- bind_cols(time=km.rec$time, n.risk=km.rec$n.risk, 
                    event=km.rec$n.event, censor=km.rec$n.censor, 
                    surv=km.rec$surv, std.err=km.rec$std.err)

km.rec_tbl$strata <- c_all.rec
km.rec_tbl <- km.rec_tbl %>%
  separate(col = strata, into = c("Cohort", "Received_RT"),sep = ", ")%>%
  mutate(Cohort = sub(pattern = "Cohort=",replacement = "", x=Cohort)) %>%
  mutate(Received_RT = sub(pattern = "Received_RT=",replacement = "", x=Received_RT)) %>%
  # mutate(Site = recode(Site, breast="Breast",breast_TN="TNBC", melanoma="Melanoma",
  #                      endometrial="Endometrium", HN="Head & Neck",lung="NSCLC")) %>%
  mutate(Outcome = "Recurrence") %>%
  mutate(Received_RT = if_else(Received_RT==0,"-RT","+RT"))

cohorts <- unique(km.rec_tbl$Cohort)

km.rec_tbl <- km.rec_tbl %>% mutate(Site = case_when(
  grepl("Breast", Cohort) ~ "Breast",
  grepl("Endo", Cohort) ~ "Endometrium",
  grepl("Glioma", Cohort) ~ "Glioma",
  grepl("Head", Cohort) ~ "Head & Neck",
  grepl("Mel", Cohort) ~ "Melanoma",
  grepl("NSCLC", Cohort) ~ "NSCLC",
  grepl("Pancreas", Cohort) ~ "Pancreas",
  grepl("TNBC", Cohort) ~ "TNBC",
  Cohort == "Pooled" ~ "Pooled"
))

km.rec_tbl<- bind_rows(km.rec_tbl,
          tibble(expand_grid(Cohort=cohorts,Received_RT = c("-RT","+RT")),surv=1,time=0) )


ggplot(km.rec_tbl) + 
  geom_step(aes(x=time,y=surv,
                # color=Site,
                color=Cohort,
                group=interaction(Cohort,Received_RT),
                linetype=Received_RT
                )) + 
  scale_x_continuous(expand=expansion(mult=0.01,add=0)) +
  scale_y_continuous(expand=expansion(mult=0.01,add=0)) +
  scale_color_manual(values=q3[unique(km.rec_tbl$Cohort)]) +
  scale_linetype_manual(values = c('42',"solid")) +
  # guides(color=guide_legend(reverse = TRUE)) +
  facet_grid(.~Received_RT) +
  xlab("Years From Diagnosis") + ylab("Recurrence-Free Proportion") + 
  theme_bw() + 
  theme(plot.title = element_text(face="bold",size=10,hjust = .5),
        legend.key.width = unit(22,units = "pt"),
        panel.spacing.x = unit(18,"pt")) +
  guides(color = guide_legend(order=1,reverse = F),
  Received_RT = guide_legend(order=2, reverse = T)) +
  coord_cartesian(xlim = c(0,10), ylim = c(0,1)) + ggtitle("KM: First Recurrence")

```

An now let's first confirm that our method of plotting KM produces the same results as the previous code. We can see that although the visualization styles are different the data appear the same for RT+ cohorts. 

```{r redo_above_recur}

surv_fit <- survfit(Surv(Time, Event) ~ Cohort, data = rt.rec)

ggsurvplot(surv_fit, data = rt.rec, size = 1,
           conf.int = FALSE, # Add confidence interval
           pval = TRUE, # Add p-value
           risk.table = FALSE, # Add risk table
           risk.table.height = 0.45, # change if you have >2 groups
           risk.table.y.text.col = T, # show colors in table listing
           ncensor.plot = FALSE,
           xlab = "Time in months", # adjust X axis label
           # break.time.by = 12, # break X axis in time intervals
           ggtheme = theme_bw(), # Change ggplot2 theme
           title = "KM: Recurrence between cohorts")

```


### Survival

```{r, fig.align="center"}
km.df <- all.os
km.os<- survfit(Surv(Time_OS,Event_OS)~ Cohort + Received_RT, data = km.df)

for (i in 1:length(km.os$strata)) {
  cur_strata = names(km.os$strata[i])
  count = unname(km.os$strata[i])
  c <- c(rep(cur_strata,count))
  if (i==1){
    c_all <- c
  } else {
    c_all <- c(c_all,c)
  }
}
km.os_tbl <- bind_cols(time=km.os$time, n.risk=km.os$n.risk, 
                    event=km.os$n.event, censor=km.os$n.censor, 
                    surv=km.os$surv, std.err=km.os$std.err)
km.os_tbl$strata = c_all
km.os_tbl <- km.os_tbl %>%
  separate(col = strata, into = c("Cohort", "Received_RT"),sep = ", ")%>%
  mutate(Cohort = sub(pattern = "Cohort=",replacement = "", x=Cohort)) %>%
  mutate(Received_RT = sub(pattern = "Received_RT=",replacement = "", x=Received_RT)) %>%
  mutate(Outcome = "Survival") %>%
  mutate(Received_RT = if_else(Received_RT==0,"-RT","+RT")) 

cohorts <- unique(km.os_tbl$Cohort)

km.os_tbl <- km.os_tbl %>% mutate(Site = case_when(
  grepl("Breast", Cohort) ~ "Breast",
  grepl("Endo", Cohort) ~ "Endometrium",
  grepl("Glioma", Cohort) ~ "Glioma",
  grepl("Head", Cohort) ~ "Head & Neck",
  grepl("Mel", Cohort) ~ "Melanoma",
  grepl("NSCLC", Cohort) ~ "NSCLC",
  grepl("Pancreas", Cohort) ~ "Pancreas",
  grepl("TNBC", Cohort) ~ "TNBC",
  Cohort == "Pooled" ~ "Pooled"
))

km.os_tbl<- bind_rows(km.os_tbl,
          tibble(expand_grid(Cohort=cohorts,Received_RT = c("-RT","+RT")),surv=1,time=0) )

ggplot(km.os_tbl) + 
  geom_step(aes(x=time,y=surv,
                group=interaction(Cohort,Received_RT),
                color=Cohort, 
                linetype=Received_RT)) + 
  # geom_point(data=filter(km_tbl, censor=="1"), 
  #            aes(x=time,y=surv,color=Site), shape=3, size=.4) +
  scale_x_continuous(expand=expansion(mult=0,add=0), limits=c(0,10)) +
  scale_y_continuous(expand=expansion(mult=0.01,add=0), limits=c(0,1)) +
  scale_color_manual(values=q3[unique(km.os_tbl$Cohort)]) +
  scale_linetype_manual(values = c('42',"solid")) +
   facet_grid(.~Received_RT) +
  guides(color=guide_legend(reverse = TRUE)) +
  xlab("Years From Diagnosis") + ylab("Overall Survival") + 
  theme_bw() + 
  theme(plot.title = element_text(face="bold",size=10,hjust = .5), 
        legend.key.width = unit(22,units = "pt"),
        panel.spacing.x = unit(18,"pt")
        ) +
  guides(color = guide_legend(order=1,reverse = F),
         Received_RT = guide_legend(order=2)) + 
  ggtitle("KM: Survival")

```

Again, we see that our new plotting method produces a similar plot comparing the RT+ groups. 
```{r redo_above_os}

surv_fit <- survfit(Surv(Time_OS, Event_OS) ~ Cohort, data = rt.os)

ggsurvplot(surv_fit, data = rt.os, size = 1,
           conf.int = FALSE, # Add confidence interval
           pval = TRUE, # Add p-value
           risk.table = FALSE, # Add risk table
           risk.table.height = 0.45, # change if you have >2 groups
           risk.table.y.text.col = T, # show colors in table listing
           ncensor.plot = FALSE,
           xlab = "Time in months", # adjust X axis label
           # break.time.by = 12, # break X axis in time intervals
           ggtheme = theme_bw(), # Change ggplot2 theme
           title = "KM: Recurrence between cohorts")


```


## KM between cohorts

### Recurrence

First, we'll clean the data to label patients in the top/bottom cohorts. In this instance, we're going to look at the 1st and 5th quintiles. 

```{r plot_recur_cohort}

gard_quint_rec <- rt.rec %>%
  group_by(Cohort) %>%
  summarise(top_cutoff = quantile(GARD, 0.80),
            bottom_cutoff = quantile(GARD, 0.20),
            median_cutoff = median(GARD))

# sanity check
test <- rt.rec %>% filter(Cohort == "Breast (Karolinksa)")
range(test$GARD)
as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (Karolinksa)", "top_cutoff"])
quantile(test$GARD, 0.8)
as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (Karolinksa)", "bottom_cutoff"])
quantile(test$GARD, 0.2)


# extract values for GARD_class cutoffs
breast_Erasmus_high <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (Erasmus)", "top_cutoff"])
breast_Erasmus_median <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (Erasmus)", "median_cutoff"])
breast_Erasmus_low <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (Erasmus)", "bottom_cutoff"])

breast_Karolinksa_high <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (Karolinksa)", "top_cutoff"])
breast_Karolinksa_median <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (Karolinksa)", "median_cutoff"])
breast_Karolinksa_low <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (Karolinksa)", "bottom_cutoff"])

breast_NKI_high <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (NKI)", "top_cutoff"])
breast_NKI_median <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (NKI)", "median_cutoff"])
breast_NKI_low <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Breast (NKI)", "bottom_cutoff"])

endometrium_TCC_high <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Endometrium (TCC)", "top_cutoff"])
endometrium_TCC_median <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Endometrium (TCC)", "median_cutoff"])
endometrium_TCC_low <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Endometrium (TCC)", "bottom_cutoff"])

hn_NKI_high <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Head & Neck (NKI)", "top_cutoff"])
hn_NKI_median <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Head & Neck (NKI)", "median_cutoff"])
hn_NKI_low <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Head & Neck (NKI)", "bottom_cutoff"])

melanoma_TCC_high <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Melanoma (TCC)", "top_cutoff"])
melanoma_TCC_median <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Melanoma (TCC)", "median_cutoff"])
melanoma_TCC_low <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="Melanoma (TCC)", "bottom_cutoff"])

nsclc_MCC_high <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="NSCLC (MCC)", "top_cutoff"])
nsclc_MCC_median <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="NSCLC (MCC)", "median_cutoff"])
nsclc_MCC_low <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="NSCLC (MCC)", "bottom_cutoff"])

tnbc_MCC_high <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="TNBC (MCC)", "top_cutoff"])
tnbc_MCC_median <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="TNBC (MCC)", "median_cutoff"])
tnbc_MCC_low <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="TNBC (MCC)", "bottom_cutoff"])

tnbc_NKI_high <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="TNBC (NKI)", "top_cutoff"])
tnbc_NKI_median <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="TNBC (NKI)", "median_cutoff"])
tnbc_NKI_low <- as.numeric(gard_quint_rec[gard_quint_rec$Cohort=="TNBC (NKI)", "bottom_cutoff"])


# assign class based on cutoffs
km_data_rec_grp <- rt.rec %>%
  mutate(GARD_class = case_when(Cohort == "Breast (Erasmus)" & GARD > breast_Erasmus_high ~ "GARD_High",
                                Cohort == "Breast (Erasmus)" & GARD < breast_Erasmus_low ~ "GARD_Low",
                                Cohort == "Breast (Karolinksa)" & GARD > breast_Karolinksa_high ~ "GARD_High",
                                Cohort == "Breast (Karolinksa)" & GARD < breast_Karolinksa_low ~ "GARD_Low",
                                Cohort == "Breast (NKI)" & GARD > breast_NKI_high ~ "GARD_High",
                                Cohort == "Breast (NKI)" & GARD < breast_NKI_low ~ "GARD_Low",
                                Cohort == "Endometrium (TCC)" & GARD > endometrium_TCC_high ~ "GARD_High",
                                Cohort == "Endometrium (TCC)" & GARD < endometrium_TCC_low ~ "GARD_Low",
                                Cohort == "Head & Neck (NKI)" & GARD > hn_NKI_high ~ "GARD_High",
                                Cohort == "Head & Neck (NKI)" & GARD < hn_NKI_low ~ "GARD_Low",
                                Cohort == "Melanoma (TCC)" & GARD > melanoma_TCC_high ~ "GARD_High",
                                Cohort == "Melanoma (TCC)" & GARD < melanoma_TCC_low ~ "GARD_Low",
                                Cohort == "NSCLC (MCC)" & GARD > nsclc_MCC_high ~ "GARD_High",
                                Cohort == "NSCLC (MCC)" & GARD < nsclc_MCC_low ~ "GARD_Low",
                                Cohort == "TNBC (MCC)" & GARD > tnbc_MCC_high ~ "GARD_High",
                                Cohort == "TNBC (MCC)" & GARD < tnbc_MCC_low ~ "GARD_Low",
                                Cohort == "TNBC (NKI)" & GARD > tnbc_NKI_high ~ "GARD_High",
                                Cohort == "TNBC (NKI)" & GARD < tnbc_NKI_low ~ "GARD_Low")) %>%
  drop_na(GARD_class)

# test single cutoff value
km_data_rec_grp <- rt.rec %>%
  mutate(GARD_class = case_when(Cohort == "Breast (Erasmus)" & GARD >= breast_Erasmus_median ~ "GARD_High",
                                Cohort == "Breast (Erasmus)" & GARD < breast_Erasmus_median ~ "GARD_Low",
                                Cohort == "Breast (Karolinksa)" & GARD >= breast_Karolinksa_median ~ "GARD_High",
                                Cohort == "Breast (Karolinksa)" & GARD < breast_Karolinksa_median ~ "GARD_Low",
                                Cohort == "Breast (NKI)" & GARD >= breast_NKI_median ~ "GARD_High",
                                Cohort == "Breast (NKI)" & GARD < breast_NKI_median ~ "GARD_Low",
                                Cohort == "Endometrium (TCC)" & GARD >= endometrium_TCC_median ~ "GARD_High",
                                Cohort == "Endometrium (TCC)" & GARD < endometrium_TCC_median ~ "GARD_Low",
                                Cohort == "Head & Neck (NKI)" & GARD >= hn_NKI_median ~ "GARD_High",
                                Cohort == "Head & Neck (NKI)" & GARD < hn_NKI_median ~ "GARD_Low",
                                Cohort == "Melanoma (TCC)" & GARD >= melanoma_TCC_median ~ "GARD_High",
                                Cohort == "Melanoma (TCC)" & GARD < melanoma_TCC_median ~ "GARD_Low",
                                Cohort == "NSCLC (MCC)" & GARD >= nsclc_MCC_median ~ "GARD_High",
                                Cohort == "NSCLC (MCC)" & GARD < nsclc_MCC_median ~ "GARD_Low",
                                Cohort == "TNBC (MCC)" & GARD >= tnbc_MCC_median ~ "GARD_High",
                                Cohort == "TNBC (MCC)" & GARD < tnbc_MCC_median ~ "GARD_Low",
                                Cohort == "TNBC (NKI)" & GARD >= tnbc_NKI_median ~ "GARD_High",
                                Cohort == "TNBC (NKI)" & GARD < tnbc_NKI_median ~ "GARD_Low")) %>%
  drop_na(GARD_class)

```


Now, we'll fit our survival model and compare between cohorts. 
```{r fit_recur_cohort_KM}

surv_fit <- survfit(Surv(Time, Event) ~ GARD_class, data = km_data_rec_grp)

ggsurvplot(surv_fit, data = km_data_rec_grp, size = 1,
           conf.int = FALSE, # Add confidence interval
           pval = TRUE, # Add p-value
           risk.table = FALSE, # Add risk table
           risk.table.height = 0.45, # change if you have >2 groups
           risk.table.y.text.col = T, # show colors in table listing
           ncensor.plot = FALSE,
           facet.by = c("Cohort"),
           xlab = "Time in months", # adjust X axis label
           # break.time.by = 12, # break X axis in time intervals
           ggtheme = theme_bw(), # Change ggplot2 theme
           title = "KM: Recurrence between cohorts separated by cutoff of 21")


```

### Survival

And, here, we'll repeat the above with OS instead of first recurrence as the event of interest. 
```{r class_os_KM}

gard_quint_os <- rt.os %>%
  group_by(Cohort) %>%
  summarise(top_cutoff = quantile(GARD, 0.80),
            bottom_cutoff = quantile(GARD, 0.20))


# sanity check
test <- rt.os %>% filter(Cohort == "Glioma (TCGA)")
range(test$GARD)
as.numeric(gard_quint_os[gard_quint_os$Cohort=="Glioma (TCGA)", "top_cutoff"])
quantile(test$GARD, 0.8)
as.numeric(gard_quint_os[gard_quint_os$Cohort=="Glioma (TCGA)", "bottom_cutoff"])
quantile(test$GARD, 0.2)

# extract values for GARD_class cutoffs
endometrium_TCC_high <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="Endometrium (TCC)", "top_cutoff"])
endometrium_TCC_low <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="Endometrium (TCC)", "bottom_cutoff"])
glioma_TCGA_high <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="Glioma (TCGA)", "top_cutoff"])
glioma_TCGA_low <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="Glioma (TCGA)", "bottom_cutoff"])
melanoma_TCC_high <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="Melanoma (TCC)", "top_cutoff"])
melanoma_TCC_low <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="Melanoma (TCC)", "bottom_cutoff"])
nsclc_MCC_high <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="NSCLC (MCC)", "top_cutoff"])
nsclc_MCC_low <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="NSCLC (MCC)", "bottom_cutoff"])
pancreas_TCC_high <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="Pancreas (TCC)", "top_cutoff"])
pancreas_TCC_low <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="Pancreas (TCC)", "bottom_cutoff"])
tnbc_MCC_high <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="TNBC (MCC)", "top_cutoff"])
tnbc_MCC_low <- as.numeric(gard_quint_os[gard_quint_os$Cohort=="TNBC (MCC)", "bottom_cutoff"])


# assign class based on cutoffs
km_data_os_grp <- rt.os %>%
  mutate(GARD_class = case_when(Cohort == "Endometrium (TCC)" & GARD > endometrium_TCC_high ~ "GARD_High",
                                Cohort == "Endometrium (TCC)" & GARD < endometrium_TCC_low ~ "GARD_Low",
                                Cohort == "Glioma (TCGA)" & GARD > glioma_TCGA_high ~ "GARD_High",
                                Cohort == "Glioma (TCGA)" & GARD < glioma_TCGA_low ~ "GARD_Low",
                                Cohort == "Melanoma (TCC)" & GARD > melanoma_TCC_high ~ "GARD_High",
                                Cohort == "Melanoma (TCC)" & GARD < melanoma_TCC_low ~ "GARD_Low",
                                Cohort == "NSCLC (MCC)" & GARD > nsclc_MCC_high ~ "GARD_High",
                                Cohort == "NSCLC (MCC)" & GARD < nsclc_MCC_low ~ "GARD_Low",
                                Cohort == "Pancreas (TCC)" & GARD > pancreas_TCC_high ~ "GARD_High",
                                Cohort == "Pancreas (TCC)" & GARD < pancreas_TCC_low ~ "GARD_Low",
                                Cohort == "TNBC (MCC)" & GARD > tnbc_MCC_high ~ "GARD_High",
                                Cohort == "TNBC (MCC)" & GARD < tnbc_MCC_low ~ "GARD_Low")) %>%
  drop_na(GARD_class)


# test based on predefined cutoff value 

km_data_os_grp <- rt.os %>%
  mutate(GARD_class = case_when(Cohort == "Endometrium (TCC)" & GARD > 21 ~ "GARD_High",
                                Cohort == "Endometrium (TCC)" & GARD < 21 ~ "GARD_Low",
                                Cohort == "Glioma (TCGA)" & GARD > 21 ~ "GARD_High",
                                Cohort == "Glioma (TCGA)" & GARD < 21 ~ "GARD_Low",
                                Cohort == "Melanoma (TCC)" & GARD > 21 ~ "GARD_High",
                                Cohort == "Melanoma (TCC)" & GARD < 21 ~ "GARD_Low",
                                Cohort == "NSCLC (MCC)" & GARD > 21 ~ "GARD_High",
                                Cohort == "NSCLC (MCC)" & GARD < 21 ~ "GARD_Low",
                                Cohort == "Pancreas (TCC)" & GARD > 21 ~ "GARD_High",
                                Cohort == "Pancreas (TCC)" & GARD < 21 ~ "GARD_Low",
                                Cohort == "TNBC (MCC)" & GARD > 21 ~ "GARD_High",
                                Cohort == "TNBC (MCC)" & GARD < 21 ~ "GARD_Low")) %>%
  drop_na(GARD_class)

```


```{r fit_OS_cohort_km}
surv_fit <- survfit(Surv(Time_OS, Event_OS) ~ GARD_class, data = km_data_os_grp)

ggsurvplot(surv_fit, data = km_data_os_grp, size = 1,
           conf.int = FALSE, # Add confidence interval
           pval = TRUE, # Add p-value
           risk.table = FALSE, # Add risk table
           risk.table.height = 0.45, # change if you have >2 groups
           risk.table.y.text.col = T, # show colors in table listing
           ncensor.plot = FALSE,
           facet.by = c("Cohort"),
           xlab = "Time in months", # adjust X axis label
           # break.time.by = 12, # break X axis in time intervals
           ggtheme = theme_bw(), # Change ggplot2 theme
           title = "KM: Overall Survival between cohorts separated by cutoff of 21")
```